---
title: "Earth Engine App Infos"
author: "Posted by Philipp Gärtner, `r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: yes
    fig_width: 9
    fig_caption: yes
    keep_md: yes
---

**[Earth Engine Apps](https://developers.google.com/earth-engine/guides/apps) are dynamic, shareable user interfaces for Earth Engine analyses. The [ee-appshot](https://github.com/samapriya/ee-appshot) repository from [Samapriya Roy](https://twitter.com/samapriyaroy) creates a weekly snapshot of available Earth Engine Apps and provides their URL's and script source codes.**

**The ee-appshot-info repository analyzes the script source codes and aggregates information on how creators use Earth Engine, what data they load, and which areas they are interested in.**

```{r setup, include=FALSE}

knitr::opts_chunk$set(fig.align="center")
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

knitr::opts_chunk$set(
  fig.path = "man/figures/README-"
)

```

```{r load-packages}

# LOAD PACKAGES ----------
library(tidyverse) 
#library(stringr)
#library(dplyr)
#library(scales)
library(treemapify)
library(patchwork)
#library(ggplot2)
#library(maps)
library(downloader) 
library(rnaturalearth)
library(rnaturalearthdata)

```

```{r}

unlink("zip", recursive=TRUE) # delete zip folder
unlink("unzipped", recursive=TRUE) # delete zip folder
unlink("data", recursive=TRUE) # delete zip folder

```




```{r}

if (!dir.exists("zip")) {dir.create("zip")}

download(url = "https://github.com/samapriya/ee-appshot/archive/refs/heads/main.zip",
         dest="zip/ee-appshot.zip", 
         mode="wb") 

```



```{r unzip}

files  <- list.files(path = "zip")

if (!dir.exists("unzipped")) {dir.create("unzipped")}

for (i in files) {
  unzip(paste0("zip/",i), exdir="unzipped")
}

unlink("zip", recursive=TRUE) # delete zip folder
```

```{r include=FALSE}

if (!dir.exists("data")) {dir.create("data")}

file.copy(from = "unzipped/ee-appshot-main/ee-appshot.zip", to = "data", overwrite = TRUE)

unlink("unzipped", recursive=TRUE) # delete zip folder

```

```{r }

files  <- list.files(path = "data")

if (!dir.exists("data/unzipped")) {dir.create("data/unzipped")}

for (i in files) {
  unzip(zipfile = paste0("data/",i), exdir="data/unzipped")
}

```



```{r create-list-of-files, include=FALSE}

# RENAME .js to .txt -------
files <- list.files(path="data/unzipped", 
                    recursive = TRUE,
                    full.names = TRUE,
                    pattern="*.js")
newfiles <- gsub(".js$", ".txt", files)
file.rename(files, newfiles)


list_of_files <- list.files(path = "data/unzipped", 
                            recursive = TRUE,
                            full.names = TRUE,
                            pattern = "\\.txt$")

```


```{r create-empty-dataframes}

# ee_data ----
ee_data    <- data.frame(V1 = NA, V2 = NA, type = NA, path = NA)
ee_lat_lon <- data.frame(V1 = NA, V2 = NA, path = NA)
ee_req     <- data.frame(V1 = NA, V2 = NA, path = NA)
script_length <- data.frame(length = NA, path = NA)

```

```{r for_loop}

for (i in 1:length(list_of_files)){
  
  #print(i)
  #print(list_of_files[i])
  
  script    <- readLines(list_of_files[i], warn = FALSE)     # read .txt file
  script_df <- as.data.frame(script)                         # convert to df
  
  ee <- script_df %>%
    mutate_if(is.character, trimws) %>%                      # trim whitespaces
    filter(str_detect(script, "^\\//", negate = TRUE)) %>%   # delete oneline comments
    filter(str_detect(script, "^\\/", negate = TRUE))        # delete multiline comments
  
  script_length[i,1] <- nrow(ee)
  script_length[i,2] <- list_of_files[i]
 
  ## find ee.Image(' xxx ') ----
  ee_image <- str_match(ee$script, "ee.Image\\('\\s*(.*?)\\s*\\'")
  ee_image <- as.data.frame(ee_image)
  
  ## find ee.Image(" xxx ") ----
  ee_image2 <- str_match(ee$script, "ee.Image\\(\\\"\\s*(.*?)\\s*\\\"")
  ee_image2 <- as.data.frame(ee_image2)
  
  ee_images <- rbind(ee_image, ee_image2)
  
  ee_images <- ee_images %>%
    filter(if_all(everything(), ~ !is.na(.x))) %>%            # delete NA's
    filter(str_detect(V2, "^users", negate = TRUE)) %>%       # filter out specific user Images
    mutate(
      type = c("ee_Image"), path = list_of_files[i])
  
  ##############################################################################
  
  ## find ee.ImageCollection(' xxx ') ----
  ee_image_collection <- str_match(ee$script, "ee.ImageCollection\\('\\s*(.*?)\\s*\\'")
  ee_image_collection <- as.data.frame(ee_image_collection)
  
  ## find ee.ImageCollection(" xxx ") ----
  ee_image_collection2 <- str_match(ee$script, "ee.ImageCollection\\(\\\"\\s*(.*?)\\s*\\\"")
  ee_image_collection2 <- as.data.frame(ee_image_collection2)
  
  ee_imagecollections <- rbind(ee_image_collection, ee_image_collection2)
  
  ee_imagecollections <- ee_imagecollections %>%
    filter(if_all(everything(), ~ !is.na(.x))) %>%
    filter(str_detect(V2, "^users", negate = TRUE)) %>%
    mutate(type = c("ee_ImageCollection"),path = list_of_files[i])
  
  ##############################################################################
  
  ## find ee.FeatureCollection(' xxx ') ----
  ee_feature_collection <- str_match(ee$script, "ee.FeatureCollection\\('\\s*(.*?)\\s*\\'")
  ee_feature_collection <- as.data.frame(ee_feature_collection)
  
  ## find ee.FeatureCollection(" xxx ") ----
  ee_feature_collection2 <- str_match(ee$script, "ee.FeatureCollection\\(\\\"\\s*(.*?)\\s*\\\"")
  ee_feature_collection2 <- as.data.frame(ee_feature_collection2)
  
  ee_feature_collections <- rbind(ee_feature_collection, ee_feature_collection2)
  
  ee_feature_collections <- ee_feature_collections %>%
    filter(if_all(everything(), ~ !is.na(.x))) %>%
    filter(str_detect(V2, "^users", negate = TRUE)) %>%
    filter(str_detect(V2, "^ft\\:", negate = TRUE)) %>%
    mutate(type = c("ee_FeatureCollection"),path = list_of_files[i])
  
  ee_data <- rbind(ee_data, ee_images, ee_imagecollections, ee_feature_collections, fill=TRUE)
  
  ###
  
  ee_location <- str_match(ee$script, "setCenter\\(\\s*(.*?)\\s*\\)")
  ee_location <- as.data.frame(ee_location)
  
  ee_location <- ee_location %>%
    filter(if_all(everything(), ~ !is.na(.x))) %>%
    mutate(path = list_of_files[i])
  
  ee_lat_lon <- rbind(ee_lat_lon, ee_location, fill=TRUE)
  
  ###
  
  ee_require <- str_match(ee$script, "require\\(\\s*(.*?)\\s*\\)")
  ee_require <- as.data.frame(ee_require)
  
  ee_require <- ee_require %>%
    filter(if_all(everything(), ~ !is.na(.x))) %>%
    mutate(path = list_of_files[i])
  
  ee_req <- rbind(ee_req, ee_require, fill=TRUE)
} 

```

# App Code Length

Currently, the repository contains **`r length(list_of_files)`** scripts with a median of **`r round(median(script_length$length),0)`** lines of code (comments excluded). 

```{r plot-script-length}

#xx <- data.frame(value = script_length)
script_length$group <- 1
script_length$type <- "a"

p1 <- ggplot(script_length, aes(x = group, fill = type, y = length)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    width = .12, 
    alpha = .5,
    color = "#4B985A",
    outlier.shape = NA
  ) + 
  geom_point(
    size = 1.3,
    alpha = .15,
    position = position_jitter(seed = 1, width = .05),
    fill = alpha("#4B985A", 0.2), 
    color = "#4B985A") + 
  scale_y_log10() +
  coord_flip() + 
  scale_fill_manual(values=c("#4B985A")) +
  theme_minimal() +
  annotate("text", y = median(script_length$length), x = 1.14, 
           color = "#404040",label = paste0(median(script_length$length), " Lines"), size = 3.5) + 
  labs(x = "", y = "Code Length",
       caption = paste0("\n","Scripts: ",nrow(script_length)," • Last Updated: ", Sys.Date(),"\n", "Visualization: Philippp Gärtner (@Mixed_Pixels) • Design Inspiration: Cédric Scherer")
       ) +
  theme(legend.position = "none",
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.text.y = element_blank(),
        plot.caption = element_text(hjust = 0.5))
  
p1  
```


```{r delete-NA-TRUE2}

ee_data <- ee_data %>%
  filter(if_all(everything(), ~ !is.na(.x))) %>% # delete NA's ----
  filter(V1 != TRUE)

#head(ee_data)
```


```{r }

original <- ee_data %>%
  group_by (type) %>%
  summarise (n=n()) %>%
  mutate(rel.freq =  scales::percent(n/sum(n), accuracy = 0.1)) %>%
  arrange(desc(n)) %>%
  mutate(type = stringr::str_replace(type, "[:punct:]", ".")) %>%
  mutate(type = as_factor(type)) 
  

barplot <-  
ggplot(original, aes(x = n, y = "a", fill = type)) + 
  geom_col(position = "fill") + 
  geom_text(aes(label = paste0(n, "\n(",rel.freq,")")), position = "fill", vjust = 1.5) +
  
  geom_text(aes(label = paste0(type)), position = "fill",
            angle=90,vjust=-2,hjust = 3,colour = "black") +
  theme_minimal() +
  scale_fill_manual(values = c("#1f72c2", "#965618","#64b273" ),
                    labels=c("ee.ImageCollection", "ee.Image", "ee.FeatureCollection") ) +
  coord_flip() +
  labs(x = "", y = "", fill = "EE object requests") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
        ) 

#barplot

```
 

 
```{r}

#ee_img <- ee_data %>% dplyr::filter(type == "ee_Image") 
ee_ic  <- ee_data %>% dplyr::filter(type == "ee_ImageCollection") 
ee_fc  <- ee_data %>% dplyr::filter(type == "ee_FeatureCollection") 

```



```{r}

get_values <- function(x, group_size, collectiontype) {
  result <-  x %>% 
    dplyr::mutate(
        item = stringr::str_to_upper(V2),
        group1 = str_split(V2, "/", simplify = TRUE, 4)[,1],
        group2 = str_split(V2, "/", simplify = TRUE, 4)[,2],
        group = paste0(as.character(group1), "/", as.character(group2), "/"),
        group = forcats::fct_lump(group, n = group_size)
        ) %>%
    dplyr::count(group, sort = TRUE) %>%
    dplyr::mutate(
      group = forcats::fct_rev(forcats::fct_inorder(group)),
      group = forcats::fct_relevel(group, "Other", after = 0)
      ) %>%   
    dplyr::mutate(
      ## set justification based on data 
      ## so that only the first label is placed inside
      place = if_else(row_number() <= 2, 1, 0),
      perc = scales::percent(n / sum(n), accuracy = .1, trim = FALSE),
      perc = paste(" ",perc, " "),
      perc = if_else(row_number() == 1, paste0(perc, "of all '", collectiontype,  "' requests"), perc)
      )
  return(result)
}

#ee_img_summary     <- get_values(ee_img, 12)
ee_img_col_summary <- get_values(ee_ic, 32, "ee.ImageCollection")
ee_fea_col_summary <- get_values(ee_fc, 16, "ee.FeatureCollection")

```

<br>

# Usage of Common Earth Engine object classes
  
ee.ImageCollections are the most important objects in Earth Engine Apps (`r original[1,2]` out of `r sum(original$n)` object requests). The image sets of *`r ee_img_col_summary[1,1]`* are the most requested, followed by *`r ee_img_col_summary[2,1]`+* and *`r ee_img_col_summary[4,1]`+*. 

For the ee.FeatureCollections, the [*`r ee_fea_col_summary[1,1]`*](https://developers.google.com/earth-engine/datasets/catalog/USDOS_LSIB_SIMPLE_2017?hl=en) dataset (Simplified International Boundary Polygons) is one of the most popular datasets. 

It is noticeable that many requested datasets do not come from the Earth Engine data catalog but from the [awesome-gee-community-datasets](https://samapriya.github.io/awesome-gee-community-datasets/) collection (recognizable by the *projects*/ label)), which is also managed by [Samapriya Roy](https://twitter.com/samapriyaroy).

```{r ee-ImageCollection, fig.width = 9, fig.height=10}

## create color palette based on input data
pal <- c("gray85",rep("gray70", length(ee_img_col_summary$group)-4), "#237fd8","#1f72c2", "#1c65ac")

plot_img_col <- ggplot(ee_img_col_summary, aes(x = n, y = group, fill = group)) +
  ## draw bars
  geom_col() +
  geom_text(aes(label = perc, hjust = place),
            ## make labels left-aligned
            size = 3.5, fontface = "bold"
  ) +
  ## reduce spacing between labels and bars
  scale_x_continuous(expand = c(.01, .01)) +
  ## add custom colors
  scale_fill_manual(values = pal, guide = "none") +
  ## change plot appearance
  theme_minimal() + 
  labs(y="",title = "ee.ImageCollection\n") + 
  ## get rid of all elements except y axis labels + adjust plot margin
  theme(axis.text.y = element_text(size = 14, hjust = 1),
        plot.title = element_text(color = "#159957", size = 22),
        plot.title.position = "plot"
        ) 

#plot_img_col

```



```{r fig.width = 9, fig.height=10}

## create color palette based on input data
pal <- c("gray85",rep("gray70", length(ee_fea_col_summary$group)-4), "#75bb82" ,"#64b273","#53a964")

plot_fea_col <- ggplot(ee_fea_col_summary, aes(x = n, y = group, fill = group)) +
  ## draw bars
  geom_col() +
  geom_text(aes(label = perc, hjust = place), size = 3.5, fontface = "bold") +
  ## reduce spacing between labels and bars
  scale_x_continuous(expand = c(.01, .01)) +
  scale_fill_manual(values = pal, guide = "none") +
  theme_minimal() + 
  labs(y="",title = "ee.FeatureCollection\n",
       caption = paste0("\n","Visualization: Philippp Gärtner (@Mixed_Pixels) • Design Inspiration: Cédric Scherer • Last Updated:", Sys.Date())) + 
  theme(axis.text.y = element_text(size = 14, hjust = 1),
        plot.title = element_text(color = "#159957", size = 22),
        plot.title.position = "plot"
        )

#plot_fea_col

```

```{r patchwork,fig.width = 10, fig.height=15}

layout <- "
BBBBC
BBBBC
BBBBC
AAAAC
AAAAC
"
plot_fea_col + plot_img_col + barplot + plot_layout(design = layout)

```

 


```{r delete-NA-TRUE}

ee_req_prep <- ee_req %>%
  filter(if_all(everything(), ~ !is.na(.x))) %>% # delete NA's ----
  filter(V1 != TRUE) %>%
  distinct() %>%
  dplyr::mutate(V2 = str_sub(V2, start = 2)) %>%
  dplyr::mutate(V2 = str_sub(V2, end = -2)) %>%
  dplyr::mutate(V2 = str_replace(V2, ".js", "")) %>%
  mutate(group = str_split(V2, "/", simplify = TRUE, 4)[,2]) 


ee_req_sum <- ee_req_prep %>%
  group_by(V2) %>%
  summarise (n=n()) %>%
  ungroup() %>%
  mutate(freq =  scales::percent(n/sum(n), accuracy = 0.1)) %>%
  mutate(group = str_split(V2, "/", simplify = TRUE, 4)[,2]) 

ee_req_sum2 <- ee_req_sum %>% filter(n > 10)


```

# Popular Script Modules

It's good practice to write [modular, reusable code](https://developers.google.com/earth-engine/guides/playground?hl=en#script-modules) that can be shared between scripts without extensive copying and pasting. In total were `r length(table(ee_req_prep$V2))` unique modules requested. The most popular modules (+10 requests) are:

```{r plot_modules_with_treemap}

ggplot(ee_req_sum2, aes(area = n, fill = group, label = paste0(V2, "\n(",n, ")",sep = ""))) + 
  geom_treemap() + 
  geom_treemap_text(colour = "white", grow = TRUE,size = 20) +
  labs(fill = "") +
  theme(legend.position = "bottom",
        plot.caption = element_text(hjust = 0.5)) +
  labs(y="",
       caption = paste0("\n","Unique Modules: ", nrow(ee_req_sum)," • Last Updated: ", Sys.Date(),"\n","Visualization: Philippp Gärtner (@Mixed_Pixels)")) 

```
```{r get_lat_lon}

ee_lat_lon2 <- ee_lat_lon %>%
  filter(if_all(everything(), ~ !is.na(.x))) %>% # delete NA's ----
  filter(V1 != TRUE) %>%
  filter(!str_detect(V2, "[:alpha:]")) %>%
   mutate(V2 = stringr::str_replace(V2, "\\(", "")) %>%#filter(!str_detect(V2, "\\\")) %>%
  dplyr::mutate(
    lon = str_split(V2, ",", simplify = TRUE, 4)[,1],
    lat = str_split(V2, ",", simplify = TRUE, 4)[,2],
    zoom = str_split(V2, ",", simplify = TRUE, 4)[,3],
    zoom = str_trim(zoom),
    zoom = round(as.numeric(zoom),0),
    lat = round(as.numeric(lat),2),
    lon = round(as.numeric(lon),2),
    group = paste0(lon, ",", lat)
)
  

zoom <- as.data.frame(table(ee_lat_lon2$zoom))

zoom$Var1 <- as.character(zoom$Var1)
zoom$Var1 <- as.integer(zoom$Var1)

xx <- as.data.frame(table(ee_lat_lon2$group))

xx <- xx %>% dplyr::mutate(
    lon = str_split(Var1, ",", simplify = TRUE, 4)[,1],
    lat = str_split(Var1, ",", simplify = TRUE, 4)[,2]
    
)

```

<br>

# Areas of Interest

[Map.setCenter(lon, lat)](https://developers.google.com/earth-engine/apidocs/map-setcenter?hl=en) centers the map view to a given coordinate pair. All Map.setCenter() requests combined result in a world map of the community areas of interests. 

```{r}
xx$lon <- as.numeric(xx$lon)
xx$lat <- as.numeric(xx$lat)

world <- ne_countries(scale = "medium", returnclass = "sf")

ggplot(data = world) +
  geom_sf(color = "white", fill = "darkgrey", size = .15) + 
  geom_point(data=xx, aes(x=lon, y=lat, size = as.numeric(xx$Freq)), 
              colour="Deep Pink", fill="Pink",pch=21, alpha=I(0.7)) +
  theme_void() +
  labs(x = "", y = "", size = "") + 
  theme(legend.position = "right",
        plot.caption = element_text(hjust = 0.5)) +
  labs(y="",
       caption = paste0("\n","Map.setCenter() calls: ", nrow(xx)," • Last Updated: ", Sys.Date(),
                        " • Visualization: Philippp Gärtner (@Mixed_Pixels)")) 
 
```